### 2019-05-22 chapter_00_open_science.tex
Reorder sentences in package management section, watchdog make directive

\subsubsection{Source code management and packaging}

@@

%\paragraphbf{Package management and software repositories} 
Python as a medium for scientific computing owes its success to a rich
ecosystem of third-party packages.  Smooth deployment of these packages
requires a package manager and a supported repository. A credible hub to upload
source codes as compressed archives or as pre-compiled binary packages was
therefore indispensable. The use of repositories started with the conception of
CPAN and CRAN repositories for the languages Perl and R in 1993 --- based on
the Comprehensive TeX Archive Network (CTAN) model for \TeX\ packages. 
%
Packaging process depends on the language used in a project, however some good
practices are universal, such as supplying documentation (a README file at the
bare minimum), keeping installation paths flexible and relying on official
build tools \citep{taschuk_ten_2017}.
%
For Python, \href{https://pypi.org}{Python Package Index (PyPI)} and
\href{https://anaconda.org}{Anaconda Cloud} are major repositories for
delivering and downloading Python packages, assisted by package managers
\pack{pip} and \pack{conda} respectively.
% Almost all major
% languages\footnote{\href{https://stackoverflow.com/questions/1693529/%
% list-of-top-repositories-by-programming-language}{https://stackoverflow.com/%
% questions/1693529/list-of-top-repositories-by-programming-language}} have
% similar repositories, with exceptions such as Fortran and Matlab.
Specialized package managers have also emerged to install scientific libraries
and compilers such as \pack{easybuild} and \pack{spack}.

\subsubsection{Continuous Integration and documentation} The need for
strong reliability of packages with increasing complexity has forced developers
to use Continuous Integration (CI).
%
When the code depends on a wide variety of external dependencies, CI is useful
to avoid \fnref{https://en.wikipedia.org/wiki/Software\_rot}{software rot},
i.e.\ to avoid obsolescence with newer versions of the compiler or its
dependencies. CI is specifically important for dynamical languages and
open-source code since it allows open-source projects to welcome contributions
from other developers while maintaining a high degree of reliability.
%
Firstly the code is \emph{built} against a known, repeatable emulation
environment on a server. Thereafter, a set of \emph{unit tests} are run for
each commit (modification of the code) to check that no bugs have been
introduced. It is also important to analyse what portion of the source code is
covered by these unit tests --- a method referred to as \emph{code coverage}.
%
Almost all CI platforms are free for open-source projects --- Travis CI,
CircleCI and Appveyor to name a few. Bitbucket and GitLab have built-in CI
implementations. Websites such as Codecov and Coveralls help to chart and
analyse the code coverage history, viewable by anyone. To use these services
one simply has to write a simple file in YAML format and authenticate the
respective services with ``read'' permissions.

New users and developers often rely on good documentation and examples to start
working with an unfamiliar piece of software. Documentation generators such as
Sphinx (for Python) and Doxygen (for many languages including C, C++, Fortran
and Python) parses in-line comments into documentations as PDF, HTML and
more \citep{lee_ten_2018}. In principle, the can be executed online and deployed
using any CI. ``Read the Docs'' has, in the recent years, become the hub for
building and hosting documentation online, thus facilitating the process for
developers and considerably improving access to information for users.
%
In the FluidDyn project we this stop at automating tests and documentation of the
latest version, but it is possible to further enhance it by adding code style
checks by running ``linters'' alongside unit tests, \emph{continuous
delivery} of package releases and preserving older versions of
documentation.

@@

% Mutual aid and knowledge sharing have been important aspects of the web since
% its inception and these practices are now of foremost importance for many
% human activities.
%
%In particular, 
Learning programming and computer science is now greatly aided by websites like
the community driven encyclopedia \href{https://www.wikipedia.org/}{Wikipedia}
or the community driven forum \href{https://stackoverflow.com/}{Stack
Overflow}, where people write thousands of questions and answers on programming
each week. IRC protocol has continued to sustain the test of time. Instant
messaging channels such as \codeinline{\#python} on Freenode facilitate
interactive discussions and a form of community-level support system --- a role
which used to be filled by mailing lists alone in the past. Alternatives such
as Riot (Matrix protocol), Gitter, and Slack have gained popularity offering
different integrations solutions on top of instant messaging.
%
Through instant messaging a geographically separated team, as in the case of
the FluidDyn project, can achieve near real-time collaboration without having
to arrange video conferences or meetings.

>>>>

\subsubsection{Source code management}

@@

\subsubsection{Package management and software repositories}

Packaging process depends on the language used in a project, however some good
practices are universal, such as licensing, supplying documentation (a README
file at the bare minimum), keeping installation paths flexible and relying on
official build tools \citep{taschuk_ten_2017}.
%
A credible repository to packages as compressed archives or as pre-compiled binary
packages is indispensable to promote code reuse and reproducibility of results.
%
The use of repositories started with the conception of
CPAN and CRAN repositories for the languages Perl and R in 1993 --- based on
the Comprehensive \TeX\ Archive Network (CTAN) model for \TeX\ packages.

Python as a medium for scientific computing owes its success to a rich
ecosystem of third-party packages. \href{https://pypi.org}{Python Package
Index (PyPI)} and \href{https://anaconda.org}{Anaconda Cloud} are major
repositories for delivering and downloading Python packages, assisted by
package managers \pack{pip} and \pack{conda} respectively. The
\fnref{https://packaging.python.org}{``Python Packaging User Guide''} is an
excellent guide to install, package and distribute projects.
% Almost all major
% languages\footnote{\href{https://stackoverflow.com/questions/1693529/%
% list-of-top-repositories-by-programming-language}{https://stackoverflow.com/%
% questions/1693529/list-of-top-repositories-by-programming-language}} have
% similar repositories, with exceptions such as Fortran and Matlab.
Specialized package managers, such as \pack{easybuild} and \pack{spack}, have
also emerged to install scientific libraries and compilers.

\subsubsection{Continuous Integration and documentation} The need for
strong reliability of packages with increasing complexity has forced developers
to use Continuous Integration (CI).
%
When the code depends on a wide variety of external dependencies, CI is useful
to avoid \fnref{https://en.wikipedia.org/wiki/Software\_rot}{software rot},
i.e.\ to avoid obsolescence with newer versions of the compiler or its
dependencies. CI is specifically important for dynamical languages and
open-source code since it allows open-source projects to welcome contributions
from other developers while maintaining a high degree of reliability.
%
Firstly the code is \emph{built} against a known, repeatable emulation
environment on a server. Thereafter, a set of \emph{unit tests} are run for
each commit (modification of the code) to check that no bugs have been
introduced. It is also important to analyse what portion of the source code is
covered by these unit tests --- a method referred to as \emph{code coverage}.
%
Almost all CI platforms are free for open-source projects --- Travis CI,
CircleCI and Appveyor to name a few. Bitbucket and GitLab have built-in CI
implementations. Websites such as Codecov and Coveralls help to chart and
analyse the code coverage history, viewable by anyone. To use these services,
one simply has to write a simple file in YAML format and authenticate the
respective services with ``read'' permissions.

New users and developers often rely on good documentation and examples to start
working with an unfamiliar piece of software. Documentation generators such as
Sphinx (for Python) and Doxygen (for many languages including C, C++, Fortran
and Python) parses in-line comments into documentations as PDF, HTML and more
\citep{lee_ten_2018}. The documentation generation can be automated online and
deployed using any CI. ``Read the Docs'' has, in the recent years, become the
hub for building and hosting documentation online, thus facilitating the
process for developers and considerably improving access to information for
users.
%
In the FluidDyn project we this stop at automating tests and documentation of
the latest version, but it is possible to further exploit the CI by adding code
style checks by running ``linters'' alongside unit tests, \emph{continuous
delivery} of package releases and preserving older versions of documentation.

@@

% Mutual aid and knowledge sharing have been important aspects of the web since
% its inception and these practices are now of foremost importance for many
% human activities.
%
%In particular, 
Learning programming and computer science is now greatly aided by websites like
the community driven encyclopaedia \href{https://www.wikipedia.org/}{Wikipedia}
or the community driven forum \href{https://stackoverflow.com/}{Stack
Overflow}, where people write thousands of questions and answers on programming
each week. IRC protocol has continued to sustain the test of time. Instant
messaging channels such as \codeinline{\#python} on Freenode facilitate
interactive discussions and a form of community-level support system --- a role
which used to be filled by mailing lists alone in the past. Alternatives such
as Riot (Matrix protocol), Gitter, and Slack have gained popularity offering
different integrations solutions on top of instant messaging.
%
Through instant messaging a geographically separated team, as in the case of
the FluidDyn project, can achieve near real-time collaboration without having
to arrange video conferences or meetings.
